Joint Training of Classification and Similarity Models for \\ Intention Detection in Task-specific Chatbot

### 2020-09-10 chatbot_classification/aaai_2021/paper.tex
modified:   paper.tex

  \begin{table}
    \begin{centering}
      \begin{tabular}{|c|c|c|c|c|}
        \hline
        F1 score & Two-stage SFC & Joint SFC & Gap & \tabularnewline
        \hline
        top-1 accuracy  & \textbf{81.50} & 81.07 & -0.47 & \tabularnewline
        top-5 accuracy & 94.01 & \textbf{94.30} & +0.29 & \tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{The average classification accuracy in stage 1 on all five dataset.}
    \label{tbe:top1_5_accuracy}
  \end{table}

@@

  \subsubsection*{Is  the  fusion of a classification model and a similarity model
  helpful?}  We  compare  the performance of joint SFC with classification based
  and similarity based baselines.

@@

  Regarding the  classification  models,  joint SFC
  outperforms    TextCNN    by   21.52   percent points,   ALBERT-xxlarge   by   7.99
  percent points, BERT-large  by  7.73 percent points, RoBERTa-base by 3.79 
  percent points, RoBERTa-large by 2.04 percents on average F1 score over 5 datasets.  
  Especially, ALBERT.xxlarge based is much more  unstable  in our senario. Thus, 
  we did multiple run with different settings, e.g.,  learning rate and batch size 
  for each dataset, and  report the  best  result  here. Comparatively,  
  RoBERTa  based is  the  most stable and has the best performance among
  all baselines and all experiment settings.

  Regarding  the  similarity  model,  as  analyzed  above,  we  only  choose the
  RoBERTa-large based as our baseline. Our joint SFC achieves the improvement of
  7.09  percents  on average. Moreover, joint SFC's result on 2 datasets without
  data  sampling  for few-shot settings, which are ITG and Amazon-670k, are also
  1.24 and 2.89 percents higher than Roberta-large classification model, and are
  6.96 and 10.77 percents higher than Roberta-large similarity model (todo:
  meaning?).

  The above analysis illustrate that suitable fusing the classification model and the 
  similarity model to make them cooperate with each other is helpful to better
  performance.

  \subsubsection*{Does joint SFC improve the classification in stage 1?}
  We  compare  the  top  1  prediction and top 5 candidate labels in stage 1 from
  two-stage SFC and joint SFC in Table \ref{tbe:top1_5_accuracy}. We should note
  that  two-stage  SFC  does not optimize the stage 1 result, which is a RoBERTa
  based classification model.

  The joint training does not improve the classification performance in stage 1,
  measured  by  top  1  accuracy.  However,  it  improves  the  quality of top 5
  candidate  class  labels.  It  can  be  explained that, a classification model
  inherently optimizes the objective loss, 0-1 error of top 1 here; the sentence
  pair similairity model in the joint SFC poses positive effect in optimizing the 
  candidate labels.

  \subsubsection*{How does training sample size influence SFC?} 
  We analyze the overall  improvement  trend  of  SFC under various few-shot
  settings, and show results in Fig. \ref{fig:trend}.

@@

  \begin{figure}[t]
    \begin{centering}
      \includegraphics[scale=0.2]{picture/improvement_trend.jpg} 
      \par
    \end{centering}
    \caption{
      Improvements with different size of training data.
    }
    \label{fig:trend}
  \end{figure}

  Especially,  in the most extreme few-shot setting with only 5 training samples
  per  class  our joint SFC achieves 4.97 percent points improvements on average
  over  RoBERTa-large classification model . This observation indicates that SFC
  is  especially  well-performed  when being applied during the initial stage of
  building  task-specific  chatbot  system,  where the amount of data sample for
  each class label is extremely scarce.

>>>>

  \begin{table}
    \begin{centering}
      \begin{tabular}{|c|c|c|c|c|}
        \hline
        F1 score & Two-stage SFC & Joint SFC & Gap & \tabularnewline
        \hline
        top-1 accuracy  & \textbf{81.50} & 81.07 & -0.47 & \tabularnewline
        top-5 accuracy & 94.01 & \textbf{94.30} & +0.29 & \tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{The average classification accuracy in percentage in stage 1 on all five dataset.}
    \label{tbe:top1_5_accuracy}
  \end{table}

@@

  \subsubsection*{Is  the  fusion of classification and similarity models
  helpful?}  We  compare  joint SFC with classification and similarity based
  baselines respectively.

@@

  Regarding the  classification  models,  joint SFC
  outperforms RoBERTa-large based, the strongest among all baselines, 
  by 2.04 percent points on average F1 score over 5 datasets.  
  Especially, ALBERT.xxlarge based is rather unstable  in this senario. Thus, 
  we did multiple run with different settings and  report the  best  ones here. 
  Comparatively,  RoBERTa  based is  the  most stable and has the best performance 
  n almost all experiment settings.

  Regarding  the  similarity  model,  as  analyzed  above,  we  only  choose the
  RoBERTa-large based implementation as our baseline. Joint SFC achieves more improvement of
  7.09  percent points  on average F1 score.

  The above analysis illustrates that suitable fusing the two kinds of models
  can take their advantages.

  \subsubsection*{Does joint SFC improve the classification in stage 1?}
  We  compare  the  top-1  prediction and top-5 candidate labels in stage 1 from
  two-stage SFC and joint SFC in Table \ref{tbe:top1_5_accuracy}. We should note
  that  two-stage  SFC  does not optimize the stage 1 model, which is a RoBERTa
  based classification model.

  The joint training does not improve the classification performance in stage 1,
  measured  by  top-1  accuracy.  However,  it  improves  the  quality of top-5
  candidate  class  labels.  It  can  be  explained that, a classification model
  inherently optimizes the objective loss, 0-1 error of top-1 here; the sentence
  pair similairity model in the joint SFC poses more positive effect in optimizing the 
  candidate labels.

  \subsubsection*{How does training sample size influence SFC?} 
  We analyze the overall  improvement  trend  of  joint SFC under various few-shot
  settings, and show results in Fig. \ref{fig:trend}.

@@

  \begin{figure}[t]
    \begin{centering}
      \includegraphics[scale=0.2]{picture/improvement_trend.jpg} 
      \par
    \end{centering}
    \caption{
      Improvements  from  Joint  SFC  over  RoBERTa  based  classification  with
      different size of training data.
    }
    \label{fig:trend}
  \end{figure}

  Especially,  in the most extreme few-shot setting with only 5 training samples
  per  class  our joint SFC achieves 4.97 percent points improvements on average
  over  RoBERTa-large classification model. This observation indicates that SFC
  is  especially  well-performed  when being applied during the initial stage of
  building  task-specific  chatbot  system,  where the amount of data sample for
  each class label is extremely scarce.
